---
title: "Rain Movies"
author: "Juli Scamardo"
date: "11/19/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Brief Note

The final video is taking a very long time to run. Additionally, to generate the basemap for that final video requires a HUC8 shapefile for Colorado. This takes a long time (approximately 1hr) to download from the internet on my computer. I included a data folder with the .gdb that has those HUC8 files so you don't need to download them. 

# Using R to make movies

While most of the visualizations made in R are `static`, with a single image 
displaying the data, one can also make videos in R. These are particularly 
useful when data has a temporal component or a third axis for display. Here
we will show a few ways to make videos using R and airport weather data 

```{r setup, include=FALSE}
library(tidyverse)
library(riem) # Precipitation data
library(animation) #movie making
library(USAboundaries) # boundary data
library(gganimate) #another way to make movies
library(lubridate)
library(ggthemes)
library(tmap)
library(sf)
library(nhdplusTools)
library(nhdR)
library(rgdal)
theme_set(theme_few()) # Set ggplot theme


knitr::opts_chunk$set(echo = TRUE)
```


# Data acquisition and reshaping


## Finding our station names

```{r}
#riem_networks() # look for network names

co_stations <- riem_stations(network = 'CO_ASOS')

#View(co_stations)

stations <- c('AFF', 'FNL')

```

## Download station data

```{r}

af_p <- riem_measures(station='AFF',date_start = '2018-10-01',
                      date_end = '2019-09-30')


fc <- riem_measures(station='FNL',date_start = '2018-10-01',
                      date_end = '2019-09-30')

View(fc)

```


## Convert both to daily

```{r}

daily_pre <- fc %>% 
  bind_rows(af_p) %>%
  mutate(date = as.Date(valid)) %>%
  group_by(date, station, lon, lat) %>%
  summarize(mean_temp = mean(tmpf,na.rm=T),
            min_temp = min(tmpf,na.rm=T),
            max_temp = max(tmpf,na.rm=T),
            daily_p = sum(p01i,na.rm=T)) %>%
  as.data.frame() %>%
  as_tibble()

```


# Making videos 

## Boring first video

```{r}

slice_p <- daily_pre %>%
  slice(1:20)

p1 <- ggplot(slice_p,aes(x=date,y=mean_temp,color=station,group=station)) + 
  geom_point() +
  labs(title = 'Date: {frame_time}', x='Date',y= 'Mean Temp (f)') + 
  transition_time(date) 


animate(p1, nframes=nrow(slice_p))


```


## Leaving a trace

```{r}
p2 <- ggplot(daily_pre,aes(x=date,y=mean_temp,color=station,group=station)) + 
  geom_line() + 
  geom_point() +
  labs(title = '100 mile temp diff', x='Date',y= 'Mean Temp (f)') + 
  transition_reveal(date) + 
  scale_color_few()


animate(p2, nframes=nrow(daily_pre))



```


## Static background plot


```{r}
usa <- us_states()

co <- usa %>%
  filter(name == 'Colorado')

sf_p <- daily_pre %>%
  st_as_sf(coords=c('lon','lat'),crs=4326)


co_elev <- elevatr::get_elev_raster(co,z=6)

tm_shape(co) + 
  tm_polygons() + 
  tm_shape(co_elev) +
  tm_raster(style='cont',palette = 'Greens') + 
  tm_shape(sf_p) + 
  tm_dots(col='station',size=3,palette=c('red','white'))

```



## Quick plot

```{r}
basemap <- tm_shape(co) + 
  tm_polygons() + 
  tm_shape(co_elev) +
  tm_raster(style='cont',
            palette = 'Greens',
            title='Elevation (m)') 

##basemap
```



```{r,eval=F}
 
full_dates <- unique(sf_p$date)

sequence <- 1:length(full_dates)
saveVideo({
  for(i in sequence){
  sub_sf <- sf_p %>%
    filter(date == full_dates[i]) %>%
    filter(daily_p > 0)
  

  if(nrow(sub_sf) > 0){
  finalmap <- basemap + tm_shape(sub_sf) + 
    tm_dots(col='station',
            size=2,
            palette = c('red','white'),
            legend.show = F) 
  }else{finalmap <- basemap}
  print(finalmap + 
          tm_layout(title=full_dates[i]) + 
          tm_add_legend(type='symbol',
                        labels=c('Air Force',
                                 'Fort Collins'),
                        col=c('red','white'),
                        title='Station'))
  }
},movie.name = 'precip_flash.mp4',
ffmpeg = 'C:/Program Files (x86)/ffmpeg/bin/ffmpeg.exe',
  interval=0.05,
  ani.width=1000,
  ani.height=800)


```


# A cooler video

## Let's see if we can see weather fronts by downloading all Colorado data and
trying to get those to flash


```{r,eval=F}

all_co <- map(co_stations$id,~riem_measures(.x,date_start='2018-10-01',
                                         date_end = '2019-09-30'))
stack_co <- do.call('rbind',all_co)

daily_p <- stack_co %>% 
  mutate(date = as.Date(valid)) %>%
  group_by(date, station, lon, lat) %>%
  summarize(mean_temp = mean(tmpf,na.rm=T),
            min_temp = min(tmpf,na.rm=T),
            max_temp = max(tmpf,na.rm=T),
            daily_p = sum(p01i,na.rm=T)) %>%
  as.data.frame() %>%
  as_tibble()

save(daily_p,file='DailyP.RData')
```

## Full state video

This code is almost identical to the above, but it uses all ASOS/AWOS precip
stations


```{r, eval=F}
load('DailyP.RData')

daily_sf <- daily_p %>%
  st_as_sf(.,coords=c('lon','lat'),crs=4326,remove=F)

full_dates <- unique(daily_sf$date)

sequence <- 1:length(full_dates)
saveVideo({
  for(i in sequence){
  sub_sf <- daily_sf %>%
    filter(date == full_dates[i]) %>%
    filter(daily_p > 0)
  

  if(nrow(sub_sf) > 0){
  finalmap <- basemap + tm_shape(sub_sf) + 
    tm_dots(title.size='Daily Precip (in)',
            col='lon',
            size='daily_p',
            style='cont',
            palette='Oranges',
            legend.show = F,
            size.max=19,
            scale=3.5,
            sizes.legend=c(0,1,5,10,20))
  }else{finalmap <- basemap}
  print(finalmap + 
          tm_layout(title=full_dates[i],
        legend.position=c(0.7,0.2)))
  }
},movie.name = 'precip_flash_all.mp4',
ffmpeg = 'C:/Program Files (x86)/ffmpeg/bin/ffmpeg.exe',
  interval=0.12,
  ani.width=1000,
  ani.height=800)



```



# In Class work.

## GGanimate

Make a video showing the correlation between Air Force Academy and Fort Collins
between the min, mean, and max temps. The movie should cover the full year. 

```{r}
## edit df to tidy temp data
tidy_daily <- daily_pre %>%
  gather(type, temp, mean_temp:max_temp)

## breaks in date for graph
dates <- as.Date(c("2018-11-01", '2019-04-01', '2019-09-01'), 
                 format = '%Y-%m-%d')

## create a ggplot of mean temp, max temp, and min temp
p3 <- ggplot(tidy_daily, aes(x = date, y = temp, 
                             color = station)) + 
  geom_point()+
  geom_line()+
  facet_wrap(~type) +
  labs(title = 'Temperature Differences: AFF and FNL',
       x='Date',
       y= 'Temp (of)') + 
  scale_color_few() + 
  scale_x_date(breaks = dates, date_labels = '%b-%y') + 
  transition_reveal(date)


## animate plot
animate(p3, 150)

```



## (Animation package)

Make a video of mean temperature differences for the full Colorado dataset
(you need to download this [here](https://www.dropbox.com/s/s7zt18elgoh8q2b/DailyP.RData?dl=0)) first.

The video should look similar to the one above, but I want you to choose your own 
style (change the background, add other layers (like counties or cities)), etc...

```{r}
## download watershed boundaries
# for downloading directly from internet (takes a long time)
#db <- nhd_load(state = "CO", dsn = "WBDHU8")

#for uploading from already downloaded file
db <- './data/WBD_National_GDB.gdb'

wbd <- st_read(db, layer = "WBDHU8")%>%
  st_cast('MULTIPOLYGON')%>%
  st_transform(crs(co_elev))

co_wbd <- wbd %>%
  select(SHAPE) %>%
  st_crop(., st_bbox(co_elev))

## create new basemap
basemap2 <- tm_shape(co) + 
  tm_polygons() +
  tm_shape(co_elev) +
  tm_raster(style='cont',
            palette = 'YlOrRd',
            contrast = c(0,0.5),
            title='Elevation (m)') +
  tm_shape(co_wbd) + 
  tm_polygons(alpha = 0.2, border.col = 'black', 
              border.alpha = 0.3)

## basemap2

## create video
load('DailyP.RData')

daily_sf <- daily_p %>%
  st_as_sf(.,coords=c('lon','lat'),crs=4326,remove=F)

full_dates <- unique(daily_sf$date)

sequence <- 1:length(full_dates)
saveVideo({
  for(i in sequence){
  sub_sf <- daily_sf %>%
    filter(date == full_dates[i]) %>%
    filter(daily_p > 0)
  

  if(nrow(sub_sf) > 0){
  finalmap <- basemap2 + tm_shape(sub_sf) + 
    tm_dots(title.size='Daily Temp (oF)',
            col='mean_temp',
            size=3,
            style='cont',
            palette='plasma',
            legend.show = TRUE)
  }else{finalmap <- basemap2}
  print(finalmap + 
          tm_layout(title=full_dates[i],
        legend.position=c(0.7,0.2)))
  }
},movie.name = 'temp_flash_all.mp4',
ffmpeg = 'C:/Program Files (x86)/ffmpeg/bin/ffmpeg.exe',
  interval=0.12,
  ani.width=1000,
  ani.height=800)

```

